Performance Insights into eBPF Step by Step

### 2014-09-05 osseu_2022/presentation.tex
Init slides

\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{booktabs}
\usepackage[scale=2]{ccicons}
\usepackage{listings}
\usepackage{marvosym}
\usepackage{color}
\usepackage{xcolor}
\usepackage[document]{ragged2e}
\usepackage[export]{adjustbox}
\usepackage{fontawesome5}
\usepackage{enumitem}
\usepackage{minted}
\usemintedstyle{tango}
\usepackage[normalem]{ulem}
\usepackage{tikz}
\usetikzlibrary{
    shapes,
    positioning,
    fit
}

@@

\def\twitter{{\faTwitter}}
\def\github{{\faGithub}}
\def\email{{\faEnvelope}}

@@

    %\vspace*{1.5em}
    \fontsize{8pt}{10}\selectfont
    {\begin{flushleft}\color{black} 12-09-2022\end{flushleft}}%

@@

\tikzset{page/.style={
    draw,
    fill=gray!30,
    rounded corners=.1cm,
}}

\tikzset{prog/.style={
    draw,
    fill=yellow!30,
    minimum height=1.0cm,
    minimum width=1.8cm,
    %rounded corners=.25cm,
}}

@@

    \begin{center}
        Current state of things
    \end{center}
\end{frame}

@@

    \vspace{2.0cm}
    \href{https://pchaigno.github.io/bpf/2021/10/20/ebpf-instruction-sets.html}
         {\color{links}\fontsize{8pt}{0}\selectfont eBPF Instruction Sets}
    \end{center}
\end{frame}
\note{
    Influences size of the program and final instructions (more workarounds for
    v1 here vs v2).  Wors with -mcpu=v1/v2/v3 -mattr=+alu32 (for me ‘probe’
    wasn’t working) via llc. With clang use -mllvm -mcpu …

@@

@@

@@

    \begin{center}
        \begin{minted}[fontsize=\Large]{c}
    struct bpf_prog_pack {
        struct list_head list;
        void *ptr;
        unsigned long bitmap[];
    };
        \end{minted}
    \end{center}
\end{frame}

\note{
    57631054fae6dcc9c892ae6310b58bbb6f6e5048
    iTLB trashing due to small progs

    Followed after frontent bound bpf progs, where iTLB contributes.
}

@@

    \begin{center}
        Which perf analysis methods\\
        could work for BPF?
    \end{center}
\end{frame}

\begin{frame}[fragile]{}
    \frametitle{Talk to compiler}

    \begin{center}
        \begin{minted}[fontsize=\Large]{shell-session}
        -Rpass=.*
        -Rpass-analysis=.*
        -Rpass-missed=.*
        \end{minted}
        \vspace{0.5cm}
        \begin{minted}[fontsize=\Large]{shell-session}
        remark: load of type i32
        not eliminated [-Rpass-missed=gvn]
        \end{minted}
    \end{center}
\end{frame}
\note{
    Global Value Numbering (GVN) pass.
}

\begin{frame}[fragile]{}
    \frametitle{Talk to compiler}

    \begin{center}
        \begin{minted}[fontsize=\Large, escapeinside=||]{c}
    static __always_inline
    int bpf_example_fn(void * |\colorbox{red!20}{\textbf{restrict}}| ctx)
        \end{minted}
    \end{center}
\end{frame}
\note{
    Within the scope of restrict-qualified variable, memory access through that
    pointer, or any other pointer based on it, is not accessed through any
    pointer not based on it.
}

\begin{frame}[fragile]{}
    \frametitle{Global kernel stats / Uptime}

\begin{frame}[fragile]{}
    \frametitle{Printk / manual instrumentation}

\begin{frame}[fragile]{}
    \frametitle{fentry \& fexit / Topdown?}

    \begin{center}
        \begin{minted}[fontsize=\Large]{shell-session}
        $ perf stat -b 5 --topdown
        \end{minted}

        \begin{columns}
            \begin{column}{0.5\textwidth}
                \begin{minted}[fontsize=\large]{c}
    SEC("fentry/XXX")
    int BPF_PROG(fentry_XXX)
    {
        //...
    }
               \end{minted}
           \end{column}

           \begin{column}{0.5\textwidth}
               \begin{minted}[fontsize=\large]{c}
    SEC("fexit/XXX")
    int BPF_PROG(fentry_XXX)
    {
        //...
    }
                \end{minted}
            \end{column}
        \end{columns}
    \end{center}
\end{frame}
\note{
    Didn't realy work with perf for whatever reason (counters were always zero,
    no events triggered). But was able to reproduce by extending bpftool with
    one topdown event.
}

    \begin{center}
        \begin{minted}[fontsize=\Large]{c}
    struct bpf_prog_pack {
        struct list_head list;
        void *ptr;
        unsigned long bitmap[];
    };
        \end{minted}
    \end{center}
\end{frame}
\note{
    57631054fae6dcc9c892ae6310b58bbb6f6e5048
    iTLB trashing due to small progs

    Followed after frontent bound bpf progs, where iTLB contributes.
}

    \vspace{1.5cm}
    \begin{center}
        \begin{tikzpicture}[
                align=center,
                node distance=0.0cm
        ]
            \path
              node[prog] (prog1) {\color{red}\faFire*}
              node[prog, right=0.2cm of prog1] (prog2) {\color{red}\faFire*}
              node[prog, right=0.2cm of prog2] (prog3) {\color{blue}\faSnowflake}

              node[prog, right=0.8cm of prog3] (prog4) {\color{red}\faFire*}
              node[prog, right=0.2cm of prog4] (prog5) {\color{blue}\faSnowflake}
              node[prog, right=0.2cm of prog5] (prog6) {$\cdots$}

              node[page,
                  inner sep=0.2cm,
                  behind path,
                  fit=(prog1)(prog2)(prog3)
              ] (page) {}

              node[page,
                  inner sep=0.2cm,
                  behind path,
                  fit=(prog4)(prog5)(prog6)
              ] (page) {};

        \end{tikzpicture}
    \end{center}
\end{frame}
\note{
    BPF progs are usually rather small (e.g. 320b), so even instruction case
    line size matter (64, and AFAIK alignment during allocation is smaller than
    that). Since many applications are actually using a chain of bpf progs via
    tail call, progs position (which would be defined by the order progs are
    loaded) could influence frontent bound overhead (due to crossing cache line
    size, or page boundaries and having mixture of hot/cold progs).
}

    \begin{center}
        \begin{minted}[fontsize=\large]{bash}
    $ perf record -e intel_pt// \
        --filter 'filter bpf_prog_9baac7ecffdb457d'

    $ perf record -e intel_pt// \
        --filter 'start 0xffffffffc1612d64'
        \end{minted}
    \end{center}
\end{frame}
\note{
    To verify perf record -a --event=mem:0xffffffffc1612d64:x

    failed to set filter "filter 0xffffffffc160f3b4/0x4050" on event intel_pt//k with 22 (Invalid argument)
}

    \begin{center}
        Modeling?
    \end{center}
\end{frame}

\begin{frame}[fragile]{}
    \frametitle{Modeling}

    \begin{center}
        \begin{minted}[fontsize=\large]{bash}
    bpf: runqslower: Use task local storage

    Replace hashtab with task local storage in runqslower.
    This improves the performance of these BPF programs.
    The following table summarizes average runtime of
    these programs, in nanoseconds:

                     task-local hash-prealloc hash
    sched_wakeup     125        340           3124
    sched_wakeup_new 2812       1510          2998
    sched_switch     151        208           991
        \end{minted}
    \end{center}
\end{frame}
\note{
}

\begin{frame}[fragile]{}
    \begin{center}
        \includegraphics[width=0.65\textwidth,center]{overhead.png}
    \end{center}
\end{frame}
\note{
}

>>>>

\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{booktabs}
\usepackage[scale=2]{ccicons}
\usepackage{listings}
\usepackage{marvosym}
\usepackage{color}
\usepackage{xcolor}
\usepackage[document]{ragged2e}
\usepackage[export]{adjustbox}
\usepackage{fontawesome}
\usepackage{enumitem}
\usepackage{minted}
\usemintedstyle{tango}
\usepackage[normalem]{ulem}
\usepackage{tikz}
\usetikzlibrary{
    shapes,
    positioning
}

@@

\def\twitter{{\FA \faTwitter}}
\def\github{{\FA \faGithub}}
\def\email{{\FA \faEnvelope}}

@@

    %\vspace*{1.5em}
    \fontsize{8pt}{10}\selectfont
    {\begin{flushleft}\color{black} 14-09-2022\end{flushleft}}%

@@

@@

    \begin{center}
        \textbf{How to measure?}
    \end{center}
\end{frame}

\begin{frame}[fragile]{}
    \frametitle{Global kernel stats}

\begin{frame}[fragile]{}
    \frametitle{Good old printk}

\begin{frame}[fragile]{}
    \frametitle{Probes before/after}

    \begin{center}
        \begin{minted}[fontsize=\Large]{shell}
    kfunc:do_syscall_64 /comm == "prog"/
    { @entry[kstack] = count(); }

    kretfunc:do_syscall_64 /comm == "prog"/
    { @return[kstack] = count(); }
        \end{minted}
    \end{center}
\end{frame}
\note{
    Pros: fast
    Cons: no ordering between progs makes it hard to find a proper point to
    attach to, maxactive, variance not relevant to the bpf prog.

    Not necessarily full syscall, maybe just before/after bpf prog.
}

\begin{frame}[fragile]{}
    \frametitle{Probes before/after}

    \begin{center}
        \begin{minted}[fontsize=\Large, escapeinside=||]{c}
  struct kretprobe {
    struct kprobe kp;
    kretprobe_handler_t handler;
    kretprobe_handler_t entry_handler;
    |\colorbox{red!20}{\textbf{int maxactive;}}|
    |\colorbox{red!20}{\textbf{int nmissed;}}|
    size_t data_size;
    struct freelist_head freelist;
    struct kretprobe_holder *rph;
  };
        \end{minted}
    \end{center}
\end{frame}
\note{
    While the probed function is executing, its return address is stored in an
    object of type kretprobe_instance.  Before calling register_kretprobe(),
    the user sets the maxactive field of the kretprobe struct to specify how
    many instances of the specified function can be probed simultaneously.
    register_kretprobe() pre-allocates the indicated number of
    kretprobe_instance objects.

mention kprobe vs kfunc, the patch and that maxactive could go away (but didn't yet).
}

\begin{frame}[fragile]{}
    \frametitle{Probes before/after}

    \begin{center}
        \begin{columns}
            \begin{column}{0.5\textwidth}
                \begin{minted}[fontsize=\large]{c}
    SEC("fentry/XXX")
    int BPF_PROG(fentry_XXX)
    {
        //...
    }
               \end{minted}
           \end{column}

           \begin{column}{0.5\textwidth}
               \begin{minted}[fontsize=\large]{c}
    SEC("fexit/XXX")
    int BPF_PROG(fentry_XXX)
    {
        //...
    }
                \end{minted}
            \end{column}
        \end{columns}
    \end{center}
\end{frame}
\note{
    Pros: fast, only target progs
    Cons: requires BTF
}

\begin{frame}[fragile]{}
    \frametitle{Probes before/after}

    \begin{center}
        \begin{minted}[fontsize=\Large]{c}
    if (!btf) {
      bpf_log(log,
        "FENTRY/FEXIT program can only be"
        "attached to another program"
        "annotated with BTF\n");
      return -EINVAL;
    }
        \end{minted}
    \end{center}
\end{frame}

\begin{frame}[fragile]{}
    \frametitle{Probes before/after}

    \begin{center}
        \begin{columns}
            \begin{column}{0.5\textwidth}
                \begin{minted}[fontsize=\large, escapeinside=||]{asm}
  |\colorbox{red!20}{\textbf{nopl   0x0(%rax,%rax,1)}}|
   xchg   %ax,%ax
   push   %rbp
   mov    %rsp,%rbp
   sub    $0x20,%rsp
               \end{minted}
           \end{column}

           \begin{column}{0.5\textwidth}
                \begin{minted}[fontsize=\large, escapeinside=||]{asm}
  |\colorbox{red!20}{\textbf{callq  0xffffffffffe0096c}}|
   xchg   %ax,%ax
   push   %rbp
   mov    %rsp,%rbp
   sub    $0x20,%rsp
                \end{minted}
            \end{column}
        \end{columns}
    \end{center}
\end{frame}
\note{
    In this way bpftool profile & perf stat -b work
}

\begin{frame}[fragile]{}
    \frametitle{Probes before/after}

    \begin{center}
        \begin{columns}
            \begin{column}{0.5\textwidth}
                \begin{minted}[fontsize=\large, escapeinside=||]{shell-session}
  $ bpftool prog profile \
      id 5 \
      duration 10 \
      cycles instructions
                \end{minted}
            \end{column}
        \end{columns}
    \end{center}
\end{frame}
\note{
    In this way bpftool profile & perf stat -b work
}

    \begin{center}
        \textbf{How to improve?}
    \end{center}
\end{frame}

@@

    \vspace{2.0cm}
    \href{https://pchaigno.github.io/bpf/2021/10/20/ebpf-instruction-sets.html}
         {\color{links}\fontsize{10pt}{0}\selectfont eBPF Instruction Sets}
    \end{center}
\end{frame}
\note{
    Influences size of the program and final instructions (more workarounds for
    v1 here vs v2).  Wors with -mcpu=v1/v2/v3 -mattr=+alu32 (for me ‘probe’
    wasn’t working) via llc. With clang use -mllvm -mcpu …

@@

    \begin{center}
        \begin{minted}[fontsize=\Large, escapeinside=||]{shell-session}
      $ llc probe.bc
          |\colorbox{red!20}{\textbf{-mcpu=v2}}|
          |\colorbox{red!20}{\textbf{-march=+alu32}}|
          -march=bpf
          -filetype=obj
          -o probe.o
        \end{minted}
        \begin{minted}[fontsize=\Large, escapeinside=||]{shell}
      # otherwise -mllvm -mcpu=...
        \end{minted}

    \end{center}
\end{frame}
\note{
    Arithmetic logic unit
}

\begin{frame}[fragile]{}
    \frametitle{BPF 2 BPF}

    \begin{center}
        \begin{minted}[fontsize=\large]{c}
    #ifndef __inline
    # define __inline                         \
       inline __attribute__((always_inline))
    #endif

    static __inline int test_bpf2bpf(void) {}
        \end{minted}

        \begin{minted}[fontsize=\large]{asm}
    # 0xffffffffc1513a68:
      nopl   0x0(%rax,%rax,1)
      xor    %eax,%eax
      push   %rbp
    # [...]
        \end{minted}
    \end{center}
\end{frame}

\begin{frame}[fragile]{}
    \frametitle{BPF 2 BPF}

    \begin{center}
        \begin{minted}[fontsize=\large]{c}
    static int test_bpf2bpf(void) { }
        \end{minted}

        \begin{minted}[fontsize=\large]{asm}
    # 0xffffffffc15b810c:
      nopl   0x0(%rax,%rax,1)
      xor    %eax,%eax
    # [...]
      callq  0x0000000000002370

    # 0xffffffffc15ba47c:
      nopl   0x0(%rax,%rax,1)
      xchg   %ax,%ax
    # [...]
        \end{minted}
    \end{center}
\end{frame}
\note{
    Works similar to bpf helpers.
    The calling convention known from BPF helper function applies to BPF to BPF
    calls just as well, meaning r1 up to r5 are for passing arguments to the
    callee and the result is returned in r0. r1 to r5 are scratch registers
    whereas r6 to r9 preserved across calls the usual way. The maximum number
    of nesting calls respectively allowed call frames is 8. A caller can pass
    pointers (e.g. to the caller’s stack frame) down to the callee, but never
    vice versa.

    BPF JIT compilers emit separate images for each function body and later fix
    up the function call addresses in the image in a final JIT pass. This has
    proven to require minimal changes to the JITs in that they can treat BPF to
    BPF calls as conventional BPF helper calls.
}

\begin{frame}[fragile]{}
    \frametitle{BPF 2 BPF}

    \begin{center}
        \begin{minted}[fontsize=\large]{c}
  if (idx && subprog[idx].has_tail_call && depth >= 256) {
    verbose(env,
      "tail_calls are not allowed when call stack"
      "of previous frames is %d bytes. Too large\n",
      depth);
    return -EACCES;
  }
        \end{minted}
    \end{center}
\end{frame}
\note{
    protect against potential stack overflow that might happen when
    bpf2bpf calls get combined with tailcalls. Limit the caller's stack
    depth for such case down to 256 so that the worst case scenario
    would result in 8k stack size (32 which is tailcall limit * 256 =
    8k).

    To get the idea what might happen, see an example:
    func1 -> sub rsp, 128
     subfunc1 -> sub rsp, 256
     tailcall1 -> add rsp, 256
      func2 -> sub rsp, 192 (total stack size = 128 + 192 = 320)
      subfunc2 -> sub rsp, 64
      subfunc22 -> sub rsp, 128
      tailcall2 -> add rsp, 128
       func3 -> sub rsp, 32 (total stack size 128 + 192 + 64 + 32 = 416)

    tailcall will unwind the current stack frame but it will not get rid
    of caller's stack as shown on the example above.
}

    \begin{center}
        \textbf{How to improve in the future?}
    \end{center}
\end{frame}

@@

    \begin{center}
        \begin{minted}[fontsize=\Large]{c}
    struct bpf_prog_pack {
        struct list_head list;
        void *ptr;
        unsigned long bitmap[];
    };

    // [...]

    ro_header = bpf_prog_pack_alloc(
                size, bpf_fill_ill_insns);
        \end{minted}
    \end{center}
\end{frame}
\note{
    57631054fae6dcc9c892ae6310b58bbb6f6e5048
    iTLB trashing due to small progs
}

@@

@@
